---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# redartR

<!-- badges: start -->
<!-- badges: end -->

`redartR` is a builder for PCoA plots from the `dartR` package. The original `dartR::gl.pcoa.plot` function builds the `ggplot` object internally and returns it to the user, but it can then sometimes be difficult to modify it to suit your needs.

This `gl.pcoa.plot.builder` function takes a slightly different appraoch, returning a object of class `ggbuilder`, which contains the data and code required to produce the `ggplot` object. You can modify that code and/or data inside the `ggbuilder` object.
Printing or plotting the `ggbuilder` object then evaluates that code and produces a `ggplot` object; printing or plotting *that* object will draw the plot to the current graphics device, as per any other `ggplot` object.

## Example
```{r example}
## load the dartR packge
library(dartR)
## and the functions from this repo
source("redartR.R")

## example plot
gl <- testset.gl
levels(pop(gl)) <- c(rep("Coast", 5), rep("Cooper", 3),rep("Coast", 5),
                     rep("MDB", 8), rep("Coast", 7), "Em.subglobosa",
                     "Em.victoriae")
pcoa <- gl.pcoa(gl, nfactors = 5)
```

The existing `gl.pcoa.plot` function:

```{r}
gl.pcoa.plot(pcoa, gl, ellipse = TRUE, p = 0.95, labels = "pop", hadjust = 1.5, vadjust = 1)
```

Construct the ggbuilder object:

```{r}
xb <- gl.pcoa.plot.builder(pcoa, gl, ellipse = TRUE, p = 0.95, labels = "pop", hadjust = 1.5, vadjust = 1)

class(xb)
```

This object contains a bunch of elements of class `ggplotter`, along with a `plot_sequence` element:

```{r}
str(xb, max.level = 2)
```

Each element of class `ggplotter` contains the data and code required to produce one part of the plot, and the `plot_sequence` elements gives the order in which those elements should be executed.

Look at one of the `ggplotter` objects:

```{r}
str(xb$points)
```

It contains a `plotfun` element (the function that will be called to plot this element) and `plotargs` (the parameters that will be passed to that function).

Plotting the `xb` object produces a ggplot object:
```{r}
xp <- plot(xb)
class(xp)
```

And printing or plotting the `xp` object produces the actual plot:
```{r}
xp
```

The whole point of this approach is that the plot can be manipulated by changing the  `ggplotter` elements. We can modify the code or data in these elements *before* calling `plot(xb)` to generate the actual `ggplot`. For example, let's replace the `ggplotter` element that controls the points:

```{r}
xb$points <- as_plotter(plotfun = "ggplot2::geom_point",
                        plotargs = list(mapping = aes_string(shape = "pop", size = 2)))
```

And now draw the modified plot:

```{r}
xb
```

Note: this is largely just a proof of concept. Only `labels` of type `"pop"`, `"ind"`, `"interactive"`, or `"plotly"` are supported at the moment.

This idea draws on a similar approach used in the [SOmap](https://github.com/AustralianAntarcticDivision/SOmap/) package.

