---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# redartR

<!-- badges: start -->
<!-- badges: end -->

`redartR` is a builder for PCoA plots from the `dartR` package. The original `dartR::gl.pcoa.plot` function builds the `ggplot` object directly, which can then be difficult to modify to suit your needs.

Instead, this `gl.pcoa.plot.builder` function returns an object of class `ggbuilder`, which contains the data and code required to produce the `ggplot` object. You can modify that code and/or data inside the `ggbuilder` object.
Printing or plotting the `ggbuilder` object then evaluates that code and produces a `ggplot` object; printing or plotting *that* object will draw the plot to the current graphics device, as per any other `ggplot` object.

## Example
```{r example}
## load the dartR packge
library(dartR)
## and the functions from this repo
source("redartR.R")

## example plot
gl <- testset.gl
levels(pop(gl)) <- c(rep("Coast", 5), rep("Cooper", 3),rep("Coast", 5),
                     rep("MDB", 8), rep("Coast", 7), "Em.subglobosa",
                     "Em.victoriae")
pcoa <- gl.pcoa(gl, nfactors = 5)

## the existing `gl.pcoa.plot` function
gl.pcoa.plot(pcoa, gl, ellipse = TRUE, p = 0.99, labels = "pop", hadjust = 1.5, vadjust = 1)


## construct the ggbuilder object
xb <- gl.pcoa.plot.builder(pcoa, gl, ellipse = TRUE, p = 0.99, labels = "pop", hadjust = 1.5, vadjust = 1)
xb <- gl.pcoa.plot.builder(pcoa, gl, labels = "pop", hadjust = 1.5, vadjust = 1)

class(xb)

## the contents of that object
str(xb)

## plotting that object produces a ggplot object
xp <- plot(xb)
class(xp)

## printing or plotting the xp object produces the actual plot
xp

## modify something in the ggbuilder object
## replace the code that plots the points
xb$points <- as_plotter(plotfun = "ggplot2::geom_point",
                        plotargs = list(mapping = aes_string(shape = "pop", size = 2)))

## and draw the modified plot
xb

```

Note: this is largely just a proof of concept. Only `labels` of type `"pop"`, `"ind"`, `"interactive"`, or `"plotly"` are supported at the moment.

